---
title: "Insurance Policy"
author: "Ted Zybin, Rohit Nair, Ibrahim Alwishah"
date:   "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## External Requirements

```{r}

# libraries

library("tidyr")
library("dplyr")
library("ggplot2")
library("mosaic")

# data
insurance_data <- read.csv("Insurance_policy.csv")
```

```{r}
knitr::kable(head(insurance_data))
```

## Research Question
How do the provided variables associate with the prediction that a customer is high value?

## Introduction
<<<<<<< HEAD
There is a belief that insurance companies use special techniques in order to maximize their profits, like: marking their clients. Agents scan customers’ portfolios, analyze their background and give them a score. Besides, any insurance company tries to put its customers in two groups: «high» and «low» value clients. A high-value customer is someone whose business will significantly impact your bottom line. They rely on your product or services more than other customers and/or bring in more business in the long run. Why Is It Important to Identify High-Value Customers?
The answer is obvious! High value customers generate more value for one’s company. High-value customers are also loyal customers. A customer retention report collated by Review42 says that:
* 65% of a company’s business comes from repeat customers
* 80% of a company’s profit is generated by 20% of its clientele
* The top 10% of your high-value customers buy 3x more than other customers
Other reasons to look for high value customers: 	
•	Knowing your high-value customers helps you to plan ahead and make better decisions about future projects.
* You can maximize their value and develop longer-lasting relationships.
* You can focus your marketing and acquisition strategies on similar customers using lookalike audiences.
* Delighting these customers can create a positive feedback loop, where they bring in more business for your company.

That is why our team decided to try to create different models that will help the insurance company to predict, whether a client is marked as «high» value or not. We received the data set from the insurance company directly. Talking about some specific things about the data set, we would like to mention that it consists almost of 50 thousands rows. Clients are not divided by gender, but the insurance company pays more attention to the marital status of their clients, their level of education and their insurance score. In previous parts of the project, my teammates and I created different models and coded different graphs to trace the correlation between the variables.  
	Another essential feature is that the insurance company will use the prediction method for clients who are older than 25 years old. This assumption is made to lower the risk of coverage and maximize its profits. 
=======
There is a belief that insurance companies use special techniques in order to maximize their profits, like: marking their clients. Agents scan customers’ portfolios, analyze their background and give them a score. Besides, any insurance company tries to put its customers in two groups: «high» and «low» value clients. A high-value customer is someone whose business will significantly impact your bottom line. They rely on your product or services more than other customers and/or bring in more business in the long run. Why Is It Important to Identify High-Value Customers?
The answer is obvious! High value customers generate more value for one’s company. High-value customers are also loyal customers. A customer retention report collated by Review42 says that:
65% of a company’s business comes from repeat customers
80% of a company’s profit is generated by 20% of its clientele
The top 10% of your high-value customers buy 3x more than other customers
Other reasons to look for high value customers: 	
•	Knowing your high-value customers helps you to plan ahead and make better decisions about future projects.
You can maximize their value and develop longer-lasting relationships.
You can focus your marketing and acquisition strategies on similar customers using lookalike audiences.
Delighting these customers can create a positive feedback loop, where they bring in more business for your company.

That is why our team decided to try to create different models that will help the insurance company to predict whether a client is marked as «high» value or not. We received the data set from the insurance company directly. Talking about some specific things about the data set, we would like to mention that it consists of almost 50 thousands rows. Clients are not divided by gender, but the insurance company pays more attention to the marital status of their clients, their level of education and their insurance score. In previous parts of the project, my teammates and I created different models and coded different graphs to trace the correlation between the variables.  
>>>>>>> main

## Visualization

## Descriptive Statistics

## Analysis

### Quantify Uncertainty

-   Bootstrap Confidence Intervals

## Discussion

## 

## References

## Appendix
