---
title: "Insurance Policy"
author: "Ted Zybin, Rohit Nair, Ibrahim Alwishah"
date:   "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## External Requirements

```{r}

# libraries

library("tidyr")
library("dplyr")
library("ggplot2")
library("mosaic")

# data
insurance_data <- read.csv("Insurance_policy.csv")
```

```{r}
knitr::kable(head(insurance_data))
```

## Research Question

How do the provided variables associate with the prediction that a
customer is high value?

## Introduction

Risk assessment is a critical aspect of the insurance industry, as it
allows companies to accurately assess the likelihood of a policyholder
filing a claim and to set premiums accordingly. By identifying high
value policyholders, insurance companies can prioritize providing them
with the best possible products and services, leading to increased
customer satisfaction and loyalty. Additionally, accurately predicting
which policyholders are likely to be low value can help the company make
informed decisions about which potential customers to accept or reject.
This can help the company avoid accepting high-risk policyholders who
are likely to file costly claims, ultimately leading to improved
financial performance. Marginal improvements in risk assessment
techniques can therefore have a significant impact on the overall
success of an insurance company.

In this study, we will explore the relationship between various
predictor variables and the target variable of whether an individual is
considered a "high value" or "low value" policyholder by an insurance
company. We have been provided with data on past policyholders,
including their age, level of education, marital status, occupation,
capital gains, hours worked per week, and an insurance score. Using this
data, we will construct a model that can accurately predict whether a
prospective policyholder will be of high or low value to the company.
While predictive accuracy is of utmost importance, we will also strive
to interpret our results in a way that makes intuitive sense. For the
purposes of this analysis, we will only consider individuals who are 25
years of age or older.

Based on the data provided, the factors of interest include:

-   The age of the prospective policyholder

-   A numerical indicator of the amount of education that a policyholder
    has \* Marital status of the policyholder

-   Occupation of the policyholder

-   Capital gains recorded on the investments of the policyholder

-   The number of hours worked per week by the policyholder

-   The proprietary insurance score of the policyholder

Furthermore, we can synthesize new variables from the existing one that
would prove useful in ultimately predicting that value of a potential
policyholder.

-   The ratio of capital gains to hours worked per week, which could
    provide insight into the policyholder's investment habits and
    financial success.

-   A variable which captures the effect of increasing age on the value
    of education

-   Use regression analysis to derive a predictive model that
    incorporates all of the existing predictor variables, which would
    allow us to identify the relative importance of each variable and
    potentially create a new, composite variable that represents the
    combined effect of all predictors.

**Insert Explanation of Models**

### Visualization

-   Identify the scientific questions we can ask about the data.
-   Create new variables & information given the existing variables in
    the dataset.
-   Write custom functions to transform data

### Descriptive Statistics

-   Present the data in different types of tables, diagrams, and find
    all descriptive stats (mean, median, mode, std.dev, etc.,)
-   3 numerical summary tables and 3 graphs

### Analysis

-   Discuss at least one bootstrap confidence interval, at least one
    hypothesis test, and two linear models trained on all the data
    (include at least one scatter-plot with a regression line). Compare
    the two linear models in terms of test mean square error (train both
    using the same subset of all the data called training data, and form
    the test mean square error using the rest of the data, called test
    data).

## Discussion + Conclusion

-   Write a clear description of the results from the previous sections.
    Do not forget to address the research questions that you posed in
    the introduction section. Make sure to pose possible questions based
    on graphs, tables, and other inferential results.

## References

-   List of all books, journals, web sites, videos referenced.
-   Use footnotes for in-line references (see sample project)

## Appendix

-   All code goes here.
